\documentclass{article}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[all]{xy}
\usepackage[pagebackref,colorlinks]{hyperref}
\usepackage{mathrsfs}

% Color comments!
\usepackage[usenames,dvipsnames]{color}
% Color comments

\newcommand{\jenn}[1]{{\color{magenta} \sf $\clubsuit\clubsuit\clubsuit$ Jenn: [#1]}}
\newcommand{\shahed}[1]{{\color{Purple} \sf $\clubsuit\clubsuit\clubsuit$ Shahed: [#1]}}

\newcommand{\sce}{\mathscr{C}^{\textsf{ex}}}
\newcommand{\scd}{\mathscr{C}}
\newcommand{\caff}{C_{\textsf{aff}}}


\theoremstyle{plain}
\newtheorem*{reftheorem}{Theorem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{problem}{Problem}
\newtheorem{question}{Question}
\newtheorem*{question*}{Question}
\newtheorem{claim}{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

% General
\renewcommand{\emptyset}{\varnothing}
\newcommand{\hra}{\hookrightarrow}
\newcommand{\righthookarrow}{\hookrightarrow}
\newcommand{\isom}{\cong}
\newcommand{\too}{\longrightarrow}
\newcommand{\isomto}{\overset{\sim}{\longrightarrow}}
\newcommand{\nto}[1]{\overset{#1}{\longrightarrow}}
\newcommand{\nsubset}{\not\subset}
\renewcommand{\phi}{\varphi}
\newcommand{\To}{\Rightarrow}
\newcommand{\ilim}{\displaystyle\lim_{\leftarrow}}
\newcommand{\dirlim}{\displaystyle\lim_{\rightarrow}}
\newcommand{\eps}{\varepsilon}
\renewcommand{\bar}[1]{\overline{#1}}
\renewcommand{\tilde}[1]{\widetilde{#1}}
\DeclareMathOperator{\car}{char}
\DeclareMathOperator{\rk}{rk}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\pgl}{PGL}
\DeclareMathOperator{\Gl}{GL}
\DeclareMathOperator{\Sl}{SL}

% Number theory
\newcommand{\Qbar}{\ensuremath{\overline{\Q}}}
\newcommand{\Kb}{\overline{K}}
\newcommand{\Fb}{\overline{F}}
\newcommand{\kb}{\overline{k}}
\newcommand{\Xbar}{\overline{X}}
\newcommand{\Cbar}{\overline{C}}
\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\F}{\ensuremath{\mathbb{F}}}
\newcommand{\fp}{\ensuremath{\mathbb{F}_p}}
\newcommand{\sm}{\ensuremath{\mathfrak{m}}}
\newcommand{\Q}{\ensuremath{\mathbb{Q}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}
\newcommand{\ok}{\mathscr{O}_K}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\inv}{inv}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\tr}{Tr}

% Algebraic geometry
\newcommand{\sA}{\ensuremath{\mathscr{A}}}
\newcommand{\sO}{\ensuremath{\mathscr{O}}}
\newcommand{\sL}{\ensuremath{\mathscr{L}}}
\newcommand{\sK}{\ensuremath{\mathscr{K}}}
\newcommand{\sF}{\ensuremath{\mathscr{F}}}
\newcommand{\A}{\ensuremath{\mathbb{A}}}
\newcommand{\Pro}{\ensuremath{\mathbb{P}}}
\newcommand{\G}{\ensuremath{\mathbb{G}}}
\newcommand{\sG}{\mathscr{G}}
\newcommand{\sX}{\mathscr{X}}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\Div}{Div}
\DeclareMathOperator{\dv}{div}
\DeclareMathOperator{\Pic}{Pic}
\DeclareMathOperator{\P0}{Pic^0}
\DeclareMathOperator{\Spec}{Spec}

\begin{document}

\title{Regular model at infinity}
\author{Jenn Park \and Shahed Sharif}
\maketitle

Let $k$ be the finite field $\F_q$ and $K = k(t)$. Let $r \geq 3$ be an integer. Let $C$ be the smooth projective curve with affine model
\[
C: xy^r = (x+1)(x+t).
\]
The purpose of this note is to compute the minimal proper regular model of $C$ at $t = \infty$, as well as the component group.

\section{Desingularization}
\label{sec:desingularization}

Let $T = \frac{1}{t}$, so that our curve has affine piece given by
\[
\caff:Txy^r = (x+1)(Tx+1).
\]
To desingularize the generic fiber, consider the affine curves
\begin{align*}
  C'&: Tvy^r = (v+1)(v+T) \\
  C''&: Tu = z^{r-1}(z+uT)(z+u) \\
  C'''&: Tw = z^{r-1}(wz+T)(wz+1).
\end{align*}
Now glue the 4 equations together via
\begin{gather*}
  v = \frac{1}{x} = \frac{z}{u} = wz \qquad y = \frac{1}{z} \\
  u = \frac{x}{y} = \frac{1}{vy} = \frac{1}{w} \qquad z = \frac{1}{y}\\
  w = \frac{y}{x} = vy = \frac{1}{u} \qquad z = \frac{1}{y}
\end{gather*}
One checks that the resulting curve is generically smooth.
\begin{remark}
  If one projectivizes $\caff$, one obtains the equation
  \[
  TXY^r = Z^{r-1}(X+Z)(TX+Z)
  \]
  in $\Pro^2_K$. The chart $C''$ corresponds to the affine patch with $Y = 1$. When setting $X = 1$, one finds that the curve is not generically smooth; it has a cusp at $[1:0:0]$. Normalizing, one obtains the equation for $C'''$. 
\end{remark}


The special fibers on these charts are as follows:
\[
\begin{array}{lll}
\caff & x + 1 = 0 & C_1 \\
C' & v(v + 1) = 0 & C_2, C_1 \\
C'' & z^r(z + u) = 0 & R_1, C_1 \\
C''' & z^rw(wz+1) = 0 & R_1, C_2, C_1.
\end{array}
\]
The third column gives names for the components of the special fiber, given in order of the factors in the equation. For example, in $C'''$, $C_2$ is given by the equation $w = 0$. The components $C_1$ and $C_2$ each have multiplicity 1, while $R_1$ has multiplicity $r$. Let $P = C_1 \cap R_1$. Note that $P$ lies in the chart $C''$, but in no others. Let $Q = C_2 \cap R_1$. Then $Q$ lies in $C'''$ but in no other chart. One checks that every point but $P$ and $Q$ is regular. Therefore we need to blow up at these two points.

\subsection{Blowing-up}
\label{sec:blowing-up}

\subsubsection{Blow-up at $P$.}
\label{sec:blow-up-P}

We first blow up at $P$, which is given by $(u,z,T) = (0,0,0)$. The blow-up is given by two charts. The first chart is computed via the change of variables
\[
z = uv \qquad T = u\tau.
\]
We obtain the arithmetic surface with equation
\[
\tau = u^{r-1}v^{r-1}(v + u\tau)(v + 1).
\]
This has special fiber given by $T = u\tau = 0$, with components
\begin{align*}
  \tilde{C_1}&: (v + 1 = \tau = 0) \\
  \tilde{R_1}&: (v = \tau = 0) \\
  D_r&: (u = \tau = 0).
\end{align*}
In the above, $\tilde{C_1}$, $\tilde{R_1}$ are the strict transforms of $C_1$, $R_1$ respectively with the same multiplicities as before. To compute the multiplicity of the third component $D_r$, we consider the ring
\[
\frac{k[u,v,\tau]_{(u,\tau)}}{(u^{r-1}v^{r-1}(v + u\tau)(v + 1) - \tau)}.
\]
This is the local ring for the subscheme $D_r$. Since $D_r$ is a prime divisor, the local ring is a discrete valuation ring $\sO_{C,D_r}$ with valuation, say, $\nu$. Then the multiplicity of $D_r$ in the special fiber is $\nu(T) = \nu(u\tau)$. But this valuation equals the length of the Artinian ring
\[
\frac{\sO_{C,D_r}}{(u\tau)}.
\]
Let $\alpha = v^{r}(v+1)$. Then the above ring equals
\[
\frac{k[u,v,\tau]_{(u,\tau)}}{(\alpha u^{r-1} - \tau, u\tau)}.
\]
The first expressions allows us to substitute $\tau = \alpha u^{r-1}$. The second expression then becomes $\alpha u^r$. Since $\nu(\alpha) = 0$, our Artinian ring is isomorphic to
\[
\frac{k[u,v]_{(u)}}{(u^r)}.
\]
Therefore the multiplicity of $D_r$ in the special fiber is $r$.

The intersection points are $P_1 = \tilde{C_1} \cap D_r$ given by $u=v+1=\tau=0$, and $P_z = \tilde{R_1} \cap D_r$ given by $u=v=\tau=0$. (Observe that $\tilde{C_1} \cap \tilde{R_1} = \emptyset$.) Writing the equation for our surface as
\[
u^{r-1}v^{r-1}(v + u\tau)(v + 1) - \tau = 0
\]
and taking the $\tau$ derivative on the left, we obtain
\[
u^rv^{r-1}(v + 1) - 1.
\]
We now observe that this expression equals $-1$ at both $P_1$ and $P_z$. Therefore both points are regular.

Now we compute the second chart. This chart is given by the change of variables
\[
u=zx \qquad T=z\pi
\]
whence we obtain the equation
\[
x\pi = z^{r-1}(1 + zx\pi)(1 + x).
\]
The special fiber is given by $T = z\pi = 0$, which gives us the components
\begin{align*}
  \tilde{C_1}&: (1 + x = \pi = 0) \\
  D_r&: (z = \pi = 0) \\
  D_1&: (z = x = 0).
\end{align*}
Observe that $D_1$ has multiplicity $1$. The intersection $\tilde{C_1} \cap D_r$ occurs when $1 + x = \pi = z = 0$. One checks that this intersection point is in fact $P_1$. The intersection $D_r \cap D_1$ occurs when $x = \pi = z = 0$. But now note that $x = \frac{1}{v}$. Therefore this intersection point is distinct from $P_2$. (Or easier: $\tilde{R_1}$ does not appear in this chart for the same reason.) We call this new point $P_3$. One checks that $P_3$ is \emph{not} regular.

In order to resolve $P_3$, we will desingularize inductively. To ease notation, write $\alpha$ for $(1 + zx\pi)(1 + x)$, so that
\begin{itemize}
    \item the chart containing $P_3$ has equation $x\pi = \alpha z^{r-1}$,
    \item $\alpha(P_3) \neq 0$, and
    \item $T = z\pi$.
\end{itemize}

We now desingularize by replacing $P_3$ with 2 charts. The first chart is given by
\[
\pi_0 = \alpha x^{r-3} z_0^{r-1}
\]
where $\pi = x \pi_0$ and $z = x z_0$. Then $T = x^2 \pi_0 z_0$ the special fiber consists of 2 components, one given by $\pi_0 = z_0 = 0$ and of multiplicity $r$, the other given by $\pi_0 = X = 0$ having multiplicity $r-1$. One verifies that the first component is $D_r$; the second component, which we call $D_{r-1}$, is exceptional; and the intersection is transverse. Furthermore, taking the $\pi_0$ derivative shows that this chart is regular.

  The second chart is given by $x_0 \pi = \alpha z^{r-2}$, where $X = x_0 z$. The special fiber then consists of the components $\pi = z = 0$ with multiplicity $r-1$ and $x_0 = z = 0$ with multiplicity $1$. One verifies that these are $D_{r-1}$ and $D_1$ respectively. The intersection point is regular if and only if $r = 3$, in which case the claim is proved. If $r > 3$, then we are in essentially the same situation as with $P_3$ except that $x$ has become $x_0$ and $r$ has been replaced with $r-1$. We can therefore iterate, thus obtaining a ``tail'' of rational curves $D_r, D_{r-1}, \dots, D_1$, each crossing transversely with adjacent curves in the list, and such that $D_i$ has multiplicity $i$ in the special fiber.


\subsubsection{Blow-up at $Q$.}
\label{sec:blow-up-Q}

Recall that $Q$ lies on the affine chart with equation
\begin{equation}
  C''':wT = z^{r-1}(wz + T)(wz + 1)\label{eq:C'''}
\end{equation}
and has coordinates $(w,z,T) = (0,0,0)$. Just as in the case of $P$, we will bdesingularize at $Q$ recursively; however, the situation is rather more complicated. To start with, we will assume that $r \geq 4$.

We start by replacing $Q$ with two charts. Chart 1 is given by the equation
\[
T_0 = b^{r-1}w^{r-2}(bw + T_0)(bw^2 + 1)
\]
where $z = bw$ and $T = T_0w$. The special fiber consists of 2 components:
\begin{itemize}
    \item $R_1: T_0 = b = 0$ with multiplicity $r$, and
    \item $R_2: T_0 = w = 0$ also with multiplicity $r$.
\end{itemize}
 The second component is exceptional. By considering the $T_0$ derivative, one sees that this chart is regular.

Chart 2 is given by the equation
\[
w_1 T_1 = z^{r-2}(w_1z + T_1)(w_1 z^2 + 1)
\]
where $w = w_1 z$ and $T = T_1 z$. The special fiber consists of three components:
\begin{itemize}
    \item $C_2: T_1 = w_1 = 0$ with multiplicity $1$,
    \item $R_2: T_1 = z = 0$ with multiplicity $r$, and
    \item $E_1: z = w_1 = 0$ with multiplicity $1$.
\end{itemize}
The third component is exceptional. The intersection point $Q_1$ has coordinates $(w_1, z, T_1) = (0,0,0)$ and, if $r \geq 3$, is \emph{not} regular.

We therefore continue by blowing-up chart 2. We do this recursively as follows. For $1 \leq i \leq r-???$, let chart $(2,i)$ be the chart given by
  \[
  w_i T_i = z^{r-i-1} (w_i z + T_i) (w_i z^{i+1} + 1).
  \]
  Let $T = T_i z^i$. Then the special fiber ($T = 0$) consists of the components
  \begin{itemize}
      \item $C_2: T_i = w_i = 0$ with multiplicity 1,
      \item $R_{i+1}: T_i = z = 0$ with multiplicity $r$, and
      \item $E_i: z = w_i = 0$ with multiplicity $i$.
  \end{itemize}
  Note that chart 2 is the same as chart $(2,1)$. One verifies that this chart is nonregular only at $Q_i: (w_i, z, T_i) = (0, 0, 0)$. We blow up $Q_i$ via 3 charts.

  The first chart is $(2,i+1)$, which we glue to $(2,i)$ away from $Q_i$ using $w_i = z w_{i+1}$ and $T_i = z T_{i+1}$. One verifies that this is consistent with the construction above.

  The second chart is $(2,i)a$ given by $w_i = w_a T_i$ and $z = z_a T_i$. (This is an abuse of notation; one should say $w_{i,a}$ and $z_{i,a}$ or some such.) Then $T = z_a^i T_i^{i+1}$. Our chart is given by
  \[
  w_a = z_a^{r-i-1} T_i^{r-i-2} (w_a z_a T_i + 1) (w_a z_a^{i+1} T_i^{i+2} + 1).
  \]
  The special fiber consists of
  \begin{itemize}
      \item $E_i: z_a = w_a = 0$ with multiplicity $i$, and
      \item $E_{i+1}: T_i = w_a = 0$ with multiplicity $i+1$.
  \end{itemize}
  By computing the $w_a$-derivative, one checks that this chart is regular.

  The third chart is $(2,i)b$ given by $z = w_i z_b$ and $T_i = w_i T_b$. Then $T = w_i^{i+1} z_b^i T_b$ and the equation is
  \[
  T_b = w_i^{r-i-2} z_b^{r-i-1} (w_i z_b + T_b) (w_i^{i+2} z_b^{i+1} + 1).
  \]
  The chart is regular (take the $T_b$-derivative) and the special fiber has components
  \begin{itemize}
      \item $R_{i+1}: T_b = z_b = 0$ with multiplicity $r$, and
      \item $R_{i+2}: T_b = w_i = 0$ with multiplicity $r$.
  \end{itemize}

\paragraph{Case $r=3$.}
\label{sec:r=3}

We desingularize by replacing $Q$ with two charts. Chart 1 is given by the equation
\[
T_0 = b^2w(bw + T_0)(bw^2 + 1)
\]
where $z = bw$ and $T = T_0w$. The special fiber consists of 3 components:
\begin{itemize}
    \item $C_z: T_0 = b = 0$ with multiplicity 3,
    \item $C_3: T_0 = w = 0$ with multiplicity 3, and
\end{itemize}
Here, $C_3$ is exceptional. The $T_0$-derivative shows that this chart is regular.

Chart 2 is given by the equation
\[
w_1 T_1 = z(w_1z + T_1)(w_1 z^2 + 1)
\]
where $w = w_1 z$ and $T = T_1 z$. The special fiber consists of three components given by
\begin{itemize}
    \item $C_2: T_1 = w_1 = 0$ with multiplicity $1$,
    \item $C_3: T_1 = z = 0$ with multiplicity $3$, and
    \item $E_1: z = w_1 = 0$ with multiplicity $1$; the exceptional component.
\end{itemize}
One checks that this chart is singular at the 3-fold intersection point $Q_1$ given by $(w_1,z,T_1) = (0, 0, 0)$. We therefore continue by blowing up at $Q_1$.

Chart 2A is given by $w_1 = z w_2$, $T_1 = z T_2$ with equation 
\[
w_2T_2 = (w_2 z + T_2) (w_2 z^3 + 1).
\]
The chart is regular with two components:
\begin{itemize}
    \item $C_2: T_2 = w_2 = 0$, and
    \item $E_3: T_2 = z = 0$ with multiplicity 3; exceptional.
\end{itemize}

Chart 2B is given by $z = w_1 z_2$, $T_1 = w_1 T_2'$, so $T = w_1^2 z_2 T_2'$; the equation is
\[
T_2' = z_2 (w_1z_2 + T_2') (w_1^3 z_2^2 + 1).
\]
Three components:
\begin{itemize}
    \item $C_3: T_2' = z_2 = 0$ with multiplicity 3.
    \item $E_3: T_2' = w_1 = 0$ with multiplicity 3.
    \item $E_2: w_1 = z_2 - 1 = 0$ with multiplicity 2.
\end{itemize}
One verifies the chart is regular.

Chart 2C is given by $w_1 = w_3 T_1$, $z = z_3 T_1$, so $T = T_1^2 z_3$ with equation
\[
w_3 = z_3 (w_3 z_3 T_1 + 1) \alpha.
\]
Two components:
\begin{itemize}
    \item $E_2: T_1 = w_3 - z_3 = 0$ with multiplicity 2, and
    \item $E_1: z_3 = w_3 = 0$ with multiplicity 1.
\end{itemize}





\paragraph{Minimality}
\label{sec:minimality}

Next we must make sure our model is minimal via Castelnuovo's criterion, which states that
\begin{theorem}[Prop.~IV.7.5 in \cite{silvermanATAEC}]\label{thm:castelnuovo}
  Let $R$ be a discrete valuation ring with algebraically closed residue field. Let $\scd$ be a regular arithmetic surface proper over $R$ whose generic fiber is a nonsingular projective curve of genus $g \geq 1$. Then $\scd$ is minimal if and only if the special fiber of $\scd$ has no divisors $E$ for which $E \isom \Pro^1$ and $E^2 = -1$.
\end{theorem}

To compute these self-intersections we use the following theorem.
\begin{theorem}[Prop.~IV.7.3 in \cite{silvermanATAEC}]\label{thm:fibral-intersect-total}
  Let $\scd$ be a regular arithmetic surface proper over a discrete valuation ring, and let $C_k$ be the special fiber. Let $D$ be a fibral divisor. Then $D \cdot C_k = 0$.
\end{theorem}

We need to check that $\scd$ is a minimal model via Theorem~\ref{thm:castelnuovo}. We have


\section{Component group}
\label{sec:component-group}



\bibliographystyle{alpha}
\bibliography{./aimbiblio}
\end{document}