\documentclass[pagesize,paper=letter]{scrartcl}
%\documentclass{article}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[all]{xy}
\usepackage[pagebackref,colorlinks]{hyperref}
\usepackage{mathrsfs}

% Color comments!
\usepackage[usenames,dvipsnames]{color}
% Color comments

\newcommand{\sce}{\mathscr{C}^{\textsf{ex}}}
\newcommand{\scd}{\mathscr{C}}
\newcommand{\caff}{C_{\textsf{aff}}}
\newcommand{\sj}{\mathscr{J}}


\theoremstyle{plain}
\newtheorem*{reftheorem}{Theorem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{problem}{Problem}
\newtheorem{question}{Question}
\newtheorem*{question*}{Question}
\newtheorem{claim}{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

% General
\renewcommand{\emptyset}{\varnothing}
\newcommand{\hra}{\hookrightarrow}
\newcommand{\righthookarrow}{\hookrightarrow}
\newcommand{\isom}{\cong}
\newcommand{\too}{\longrightarrow}
\newcommand{\isomto}{\overset{\sim}{\longrightarrow}}
\newcommand{\nto}[1]{\overset{#1}{\longrightarrow}}
\newcommand{\nsubset}{\not\subset}
\renewcommand{\phi}{\varphi}
\newcommand{\To}{\Rightarrow}
\newcommand{\ilim}{\displaystyle\lim_{\leftarrow}}
\newcommand{\dirlim}{\displaystyle\lim_{\rightarrow}}
\newcommand{\eps}{\varepsilon}
\renewcommand{\bar}[1]{\overline{#1}}
\renewcommand{\tilde}[1]{\widetilde{#1}}
\DeclareMathOperator{\car}{char}
\DeclareMathOperator{\rk}{rk}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\pgl}{PGL}
\DeclareMathOperator{\Gl}{GL}
\DeclareMathOperator{\Sl}{SL}

% Number theory
\newcommand{\Qbar}{\ensuremath{\overline{\Q}}}
\newcommand{\Kb}{\overline{K}}
\newcommand{\Fb}{\overline{F}}
\newcommand{\kb}{\overline{k}}
\newcommand{\Xbar}{\overline{X}}
\newcommand{\Cbar}{\overline{C}}
\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\F}{\ensuremath{\mathbb{F}}}
\newcommand{\fp}{\ensuremath{\mathbb{F}_p}}
\newcommand{\sm}{\ensuremath{\mathfrak{m}}}
\newcommand{\Q}{\ensuremath{\mathbb{Q}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}
\newcommand{\ok}{\mathscr{O}_K}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\inv}{inv}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\tr}{Tr}

% Algebraic geometry
\newcommand{\sA}{\ensuremath{\mathscr{A}}}
\newcommand{\sO}{\ensuremath{\mathscr{O}}}
\newcommand{\sL}{\ensuremath{\mathscr{L}}}
\newcommand{\sK}{\ensuremath{\mathscr{K}}}
\newcommand{\sF}{\ensuremath{\mathscr{F}}}
\newcommand{\A}{\ensuremath{\mathbb{A}}}
\newcommand{\Pro}{\ensuremath{\mathbb{P}}}
\newcommand{\G}{\ensuremath{\mathbb{G}}}
\newcommand{\sG}{\mathscr{G}}
\newcommand{\sX}{\mathscr{X}}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\Div}{Div}
\DeclareMathOperator{\dv}{div}
\DeclareMathOperator{\Pic}{Pic}
\DeclareMathOperator{\P0}{Pic^0}
\DeclareMathOperator{\Spec}{Spec}

\begin{document}

\title{Regular model at $t=1$}
\maketitle

Let $k$ be the finite field $\F_q$ and $K = k(t)$. Let $r \geq 3$ be an integer. Let $C$ be the smooth projective curve with affine model
\[
C: xy^r = (x+1)(x+t).
\]
The purpose of this note is to compute the minimal proper regular model of $C$ at $t = 1$, as well as the component group.

\section{Desingularization}
\label{sec:desingularization}

\subsection{Generic fiber}
\label{sec:generic-fiber}

Our first step is to substitute $x_0 = x+1$ and $t_0 = t-1$, so that our equation becomes
\[
(x_0 - 1) y^r = x_0(x_0 + t_0).
\]
This means we are concerned with the fiber $t_0 = 0$. Furthermore, there is a nonregular point at $(x_0, y, t_0) = (0, 0, 0)$. Before blowing-up, we must find a model which is generically smooth. We do this by gluing together the equations
\begin{align*}
  \caff&: (x_0 - 1) y^r = x_0(x_0 + t_0)\\
  C^\dagger&: (1-z)z = u^r(1+t_0z) \\
  C^\ddag&: v-u = vu^{r-1}(v + t_0 u)
\end{align*}
via
\begin{gather*}
  x_0 = \frac{1}{z} = \frac{v}{u} \qquad y = \frac{1}{u} \\
  u = \frac{1}{y} \qquad z = \frac{1}{x_0} \\
  v = \frac{x_0}{y} = \frac{1}{uz}.
\end{gather*}
One verifies that the resulting curve is generically smooth. (Note: the chart $C^\ddag$ is obtained by normalizing the projective closure of $\caff$ at a single cusp.) Furthermore, one can show that $C^\dag$ and $C^\ddag$ are regular.

\subsection{Blow-ups}
\label{sec:blow-ups}

It remains to resolve the singularity at $(x_0, y, t_0) = (0, 0, 0)$. To do this, we consider the family
\[
(x_iy^i - 1) y^k = x_i(x_i + t_i)
\]
with central fiber $t_iy^i = 0$; call this curve chart $(i,k)$. Observe that our $\caff$ is exactly the chart $(0,r)$. For most values of $i, k$, the chart will be nonregular at exactly one point: $(x_i, y, t_i) = (0, 0, 0)$. As we will see, a blow-up results in 3 charts, 2 of which are regular, and the third of which is chart $(i+1, k-2)$. The process terminates when chart $(i, k)$ is regular; that is, for sufficiently small $k$. This will turn out to occur when $k = 0$ or $1$. We first study properties of the charts $(i, k)$ in 4 cases:
\begin{enumerate}
    \item The initial chart $(0, k)$ with $k = r$; in particular, $k \geq 3$.
    \item The terminal chart $(i, 0)$ with $i \geq 1$.
    \item The terminal chart $(i, 1)$ with $i \geq 1$.
    \item The generic chart $(i, k)$ with $i \geq 1$, $k \geq 2$.
\end{enumerate}
Once this is done, we demonstrate the recursive blow-up procedure.

!!! Details forthcoming for all below.

\paragraph{Chart $(0, k)$ with $k \geq 3$:}
\label{sec:case-i=0}

If $i = 0$, we have $\caff$ with special fiber
\[
(x_0 - 1) y^k = x_0^2.
\]
The fiber is irreducible with multiplicity $1$, and we call it $F$. One checks that $F$ intersects itself at point $(x_0, y) = (0, 0)$. As observed earlier, the only nonregular point on $\caff$ is given by $(x_0, y, t_0) = (0, 0, 0)$.

\paragraph{Chart $(i, 0)$ with $i \geq 1$:}
\label{sec:case-k=0}

The special fiber consists of two components:
\begin{itemize}
    \item $F_{i,0}: t_i = x_i^2 - x_iy^i + 1 = 0$ with multiplicity 1; and
    \item $D_{i,0}: y = x_i^2 + x_it_i + 1 = 0$ with multiplicity $i$.
\end{itemize}
These components intersect twice transversally, at $(x_i, y, t_i) = (\pm i, 0, 0)$. This chart is regular.



\paragraph{Chart $(i, 1)$ with $i \geq 1$:}
\label{sec:case-i-1}

The special fiber consists of 3 components:
\begin{itemize}
    \item $F_{i,1}: t_i = x_i^2 - x_iy^{i+1} + y = 0$ with multiplicity 1;
    \item $D_{i,1}: y = x_i = 0$ with multiplicity $i$; and
    \item $E_{i,1}: y = x_i + t_i = 0$ with multiplicity $i$.
\end{itemize}
These components intersect each other transversally at $(x_i, y, t_i) = (0,0,0)$, and nowhere else. This chart is regular.



\paragraph{Chart $(i, k)$ with $i\geq 1$, $k\geq 2$:}
\label{sec:higher-i-k}

The special fiber consists of 3 components:
\begin{itemize}
    \item $F_{i,k}: t_i = x_i^2 - x_iy^{i+k} + y^k = 0$ with multiplicity 1;
    \item $D_{i,k}: y = x_i = 0$ with multiplicity $i$; and
    \item $E_{i,k}: y = x_i + t_i = 0$ with multiplicity $i$.
\end{itemize}
The component $F_{i,k}$ has a node at $(x_i,y,t_i) = (0,0,0)$, and the other two components also pass through this point. This intersection point is not regular, but the rest of the chart is regular.

The only nonregular charts are $(0,k)$ and the last chart above. Thus we now investigate the blow-up of these charts at $(x_i,y,t_i) = (0,0,0)$. The first chart of the blow-up is obtained via $x_{i+1} = x_i y$, $t_{i+1} = t_i y$; one verifies that the resulting chart is $(i+1, k-2)$, where the strict transform of $F_{i,k}$ is $F_{i+1, k-2}$, and the other 2 components of chart $(i + 1, k-2)$ are exceptional. There are two more charts in the blow-up, however the specifics depend on whether $k=2$ or $k \geq 3$. In all cases, as we shall see, these charts are regular.

\paragraph{Chart $(i,k)_b$:}
\label{sec:chart-i-k_b}

Our next chart, which we call $(i,k)_b$, is obtained via $y = x_ib_i$, $t_i = x_i\tau_i$, whence the central fiber is given by $t_0 = x_i^{i+1}b_i^i\tau_i$. The resulting model is
\[
(x_i^{i+1}b_i^i - 1)x_i^{k-2} b_i^k = 1 + \tau_i.
\]
The special fiber has 3 components. When $k \geq 3$, one verifies that these are
\begin{itemize}
    \item $F_{i,k}: \tau_i = (x_i^{i+1}b_i^i - 1)x^{k-2} b_i^k - 1 = 0$ with multiplicity 1;
    \item $E_{i,k}: b_i = 1+\tau_i = 0$ with multiplicity $i$; and
    \item $E_{i+1,k}: x_i = 1+\tau_i = 0$ with multiplicity $i+1$.
\end{itemize}
Note that if $i = 0$, $E_{i,k}$ has multiplicity $0$; this actually means that the component does not exist in this case. Otherwise,
the components $E_{i,k}$ and $E_{i+1,k}$ cross transversally at $(x_i, b_i, \tau_i) = (0, 0, -1)$. The component $F'_{i,k}$ does not intersect the others regardless of the value of $i$.

When $k = 2$, we get
\begin{itemize}
    \item $F_{i,2}: \tau_i = (x_i^{i+1}b_i^i - 1) b_i^2 - 1 = 0$ with multiplicity 1;
    \item $E_{i,2}: b_i = 1+\tau_i = 0$ with multiplicity $i$; and
    \item $?_{i+1,2}: x_i = 1+\tau_i+b_i^2 = 0$ with multiplicity $i+1$.
\end{itemize}
The intersection $E_{i,2} \cap ?_{i+1,2}$ is transversal at $(0,0,0)$, while $F_{i,2}$ intersects $?_{i+1,2}$ transversally at two points: $(x_i, b_i, \tau_i) = (0, \pm i, 0)$.

Independent of the value of $k$, one verifies that this chart is regular by computing the $\tau_i$-derivative.

\paragraph{Chart $(i,k)_u$:}
\label{sec:chart-i-k_u}

This last chart is obtained via $x_i = u_it_i$, $y = v_it_i$, giving the model
\[
(u_iv_i^it_i^{i+1} - 1) v_i^k t_i^{k-2} = u_i(u_i + 1).
\]
The central fiber is given by $t_0 = t_i^{i+1}v_i^i$. One checks that the special fiber when $k \geq 3$ consists of the components
\begin{itemize}
    \item $D_{i,k}: v_i = u_i = 0$ with multiplicity $i$;
    \item $E_{i,k}: v_i = u_i + 1 = 0$ with multiplicity $i$;
    \item $D_{i+1,k}: t_i = u_i = 0$ with multiplicity $i+1$; and
    \item $E_{i+1,k}: t_i = u_i + 1 = 0$ with multiplicity $i+1$.
\end{itemize}
As with the chart $(i,k)_b$, multiplicity $0$ components mean that such components are absent when $i = 0$.
The components $D_{i+1,k}$ and $D_{i,k}$ (when the latter exists) intersect transversally once at $(u_i, v_i, t_i) = (0,0,0)$, and the components $E_{i+1,k}$ and $E_{i,k}$ intersect transversally at $(-1,0,0)$. One checks that the latter intersection point is identical to the intersection point in chart $(i,k)_b$. There are no other intersections.

When $k = 2$, the components $D_{i,2}$ and $E_{i,2}$ are unchanged. There is no component $E_{i+1,2}$, and $D_{i+1,2}$ has equation $t_i = u_i^2 + u_i + v_i^2 = 0$ with the same multiplicity. It intersects each of $D_{i,2}, E_{i,2}$ once transversally.

When $k \geq 3$, the $u_i$-derivative shows that this chart is regular. When $k = 2$, the $u_i$-derivative is zero if and only if $u_i = -\frac12$. Substituting into the $v_i$-derivative shows that the chart is regular.


\subsubsection{Putting the charts together}
\label{sec:putt-charts-togeth}

We start with the chart $(0, r)$ and blow up, inductively obtaining charts of the form $(i, r-2i)$, $(i, r-2i)b$ and $(i, r-2i)u$. The procedure terminates upon reaching a chart of the form $(s, 1)$ or $(s, 0)$; thus we treat the cases of $r$ even and odd separately. Since $r$ is fixed, the value of $k$ for each case is a function of $i$. Therefore we will frequently leave off reference to $k$ in all subscripts. Notice that all components of the form $F_{\cdot}$ are glued together, and so we may leave off the subscript entirely. Lastly, the components $D_i$ and $E_i$ each have multiplicity $i$.

We first suppose that $r$ is odd. Let $s = \frac12(r-1)$. Then the resulting regular proper model consists of the 3-fold intersection of components $F$, $D_s$, and $E_s$ as in chart $(s, 1)$, plus two ``tails'': $D_s$ connected to $D_{s-1}$, which is then connected to $D_{s-2}, \dots$, terminating with $D_1$, and similarly $E_s$ connected to $E_{s-1}, \dots, E_1$.

Now suppose $r$ is even. Let $s = \frac12 r$.

\bibliographystyle{alpha}
\bibliography{./aimbiblio}
\end{document}